---
title: "Pet Parties"
output: pdf_document
date: "2022-09-27"
---

This document serves as a minimum working example of the technology I'll be using for my thesis project; I will be focusing on a $10km^2$ area in northeastern Ohio and applying dasymetric mapping techniques to it with mock demographic data.

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

library(terra)
library(dplyr)
```

```{r echo=FALSE}

# Loading in Raster
r <- project(rast('data/nlcd_data/NLCD_2019_Land_Cover_L48_20210604_bdSV3pzqsjEt13zGmeV9.tiff'), 'EPSG:4326')
# Color table containing RGB values for each type of land
colors <- read.csv('data/nlcd_color_table.csv')
weights <- cbind(colors$Code, colors$weight)

# human perception of color distance
yuv_transform <- rbind(c(0.299, 0.587, 0.114), 
                       c(-0.14713, -0.28886, 0.436), 
                       c(0.615, -0.51499, -0.10001))


yuv_color_mat <- t(apply(as.matrix(colors[, 3:5]), 1, \(x) as.numeric(yuv_transform %*% x)))


# function to find the nearest rgb value in our legend to the rgb value seen in the raster.
# takes a numeric list of 3 integers and returns the index of the row of the colors csv
# nearest to the observed value.
match_color <- function(rgb){
  yuv <- as.numeric(yuv_transform %*% as.numeric(rgb))
  ind <- which.min(rowSums(yuv_color_mat - yuv)^2)
  return(ind)
}

# set value to be active category, so values(r) returns values and as.polygons(r) has values
activeCat(r) <- 0

# add code and name of land to values df
indices <- apply(cats(r)[[1]][, 2:4], 1, \(x) match_color(x))
hexes <- apply(colors[indices, 3:5], 1, \(x) rgb(x[1]/255, x[2]/255, x[3]/255))
legend_df <- data.frame(value = cats(r)[[1]][,1],
                        code  = colors[indices, 1],
                        type  = colors[indices, 2],
                        hex   = hexes)

r_poly <- as.polygons(r)
values(r_poly) <- values(r_poly) %>%
  left_join(legend_df) %>%
  select(-value)
r_poly <- aggregate(r_poly, by='code')

```

# Initial Problem

We have an issue. The city is having a pet party at each government building, shown below. However, due to their storied rivalry, no party can accommodate both cats and dogs, so each building must either host a cat party or a dog party.

```{r echo=FALSE}
set.seed(16)

box_from_centroid <- function(centr){
  x <- centr[1]
  y <- centr[2]
  
  radius <- .005
  vertices <- cbind(
    c(x-radius, x-radius, x+radius, x+radius),
    c(y - radius, y+radius, y+radius, y-radius)
  )
  return(vect(vertices, type='polygons', crs='EPSG:4326'))
}

lons <- sample(seq(-81.93, -81.91, length.out=100), 5)
lats <- sample(seq(41.47, 41.485, length.out=100), 5)
buildings <- cbind(lons, lats)

plot(1, type='n', xlab='Longitude', ylab='Latitude',
     xlim = c(ext(r)$xmin, ext(r)$xmax), 
     ylim=c(ext(r)$ymin, ext(r)$ymax))
points(buildings, pch=20, col='purple', cex=2)
for(i in 1:nrow(buildings)){
  plot(box_from_centroid(buildings[i,]), add=TRUE, col='purple', alpha=0.15)
}
grid()
```
In order to maximize turnout, the city wants to make sure that the buildings are hosting a party that is reflective of the population densities of the pets in the surrounding area, represented by the boxes above. However, the city only has pet census data for the blocks shown below:

```{r echo=FALSE}

lons <- seq(-81.94, -81.90, length.out=5)
lats <- seq(41.465, 41.49, length.out=5)

# create empty matrix to store vertices, with first row being the geom ID
polys <- matrix(nrow=16*5, ncol=5)
colnames(polys) <- c('geom', 'part', 'x', 'y', 'hole')

polys[, 1] <- sort(rep(1:16, 5))
polys[, 2] <- 1
polys[, 5] <- 0

for(i in 1:4){
  for(j in 1:4){
    ind <- 20 * (i - 1) + 5 * (j - 1) + 1
    
    poly_lons <- c(lons[i], lons[i], lons[i+1], lons[i+1], lons[i])
    poly_lats <- c(lats[j], lats[j+1], lats[j+1], lats[j], lats[j])
    
    vertices <- cbind(poly_lons, poly_lats)
    
    polys[ind:(ind+4), 3:4] <- vertices
  }
}
d <- data.frame(id=1:16, dog_prop = runif(16))
polys <- vect(polys, 'polygons', atts=d, crs='EPSG:4326')


plot(1, xlab='Longitude', ylab='Latitude',
     xlim = c(ext(r)$xmin, ext(r)$xmax), 
     ylim=c(ext(r)$ymin, ext(r)$ymax))
plot(polys, add=TRUE)
for(i in 1:nrow(buildings)){
  plot(box_from_centroid(buildings[i,]), add=TRUE, col='purple', alpha=0.15)
}
points(buildings, pch=20, col='purple', cex=2)
```

```{r echo=FALSE}

plot(1, xlab='Longitude', ylab='Latitude',
     xlim = c(ext(r)$xmin, ext(r)$xmax), 
     ylim=c(ext(r)$ymin, ext(r)$ymax))
plot(r_poly, col = r_poly$hex, add=TRUE)
plot(polys, lwd=2, add=TRUE)
for(i in 1:nrow(buildings)){
  plot(box_from_centroid(buildings[i,]), col='purple', alpha=0.25, add=TRUE)
}
points(buildings, pch=20, col='purple', cex=2)
```
```{r echo=FALSE}
bounds <- box_from_centroid(buildings[1,])

plot(1, xlab='Longitude', ylab='Latitude',
     xlim = c(ext(bounds)$xmin, ext(bounds)$xmax), 
     ylim = c(ext(bounds)$ymin, ext(bounds)$ymax)
)

plot(r_poly, col=r_poly$hex, add=TRUE)
plot(polys, lwd=2, add=TRUE)

plot(box_from_centroid(buildings[1,]), add=TRUE, col='purple', alpha=0.4)

# outlined point hehe
points(rbind(buildings[1,], buildings[1,]), pch=20, col='black', cex=3)
points(rbind(buildings[1,], buildings[1,]), pch=20, col='purple', cex=2)

```


```{r block-areas, echo=FALSE}
# first isolate the library that the dog party we're interested in is being held at
lib <- box_from_centroid(buildings[1,])

# filter polys  to only ones containing overlap
blocks <- polys[relate(polys, lib, 'overlaps')]

# calculate areas by land code of each block
weighted_areas <- data.frame(matrix(0, nrow=nrow(r_poly), ncol=nrow(blocks)))
colnames(weighted_areas) <- blocks$id
rownames(weighted_areas) <- r_poly$code

for(i in 1:ncol(weighted_areas)){
  cropped_r <- crop(r_poly, blocks[i])
  codes <- as.character(cropped_r$code)
  areas <- expanse(cropped_r)
  
  # using land code as row name
  weighted_areas[codes, i] <- areas
}

# apply land code weights
for(row in rownames(weighted_areas)){
  weight <- weights[weights[,1] == row, 2]
  weighted_areas[row,] <- (weighted_areas[row,] * weight)
}

weighted_block_areas <- colSums(weighted_areas)

```

```{r intersection-areas, echo=FALSE}
# same process over intersections

# calculate areas by land code of each intersection
weighted_areas <- data.frame(matrix(0, nrow=nrow(r_poly), ncol=nrow(blocks)))
colnames(weighted_areas) <- blocks$id
rownames(weighted_areas) <- r_poly$code

for(i in 1:ncol(weighted_areas)){
  x <- crop(blocks[i], lib)
  cropped_r <- crop(r_poly, x)
  codes <- as.character(cropped_r$code)
  areas <- expanse(cropped_r)
  
  # using land code as row name
  weighted_areas[codes, i] <- areas
}

# apply land code weights
for(row in rownames(weighted_areas)){
  weight <- weights[weights[,1] == row, 2]
  weighted_areas[row,] <- (weighted_areas[row,] * weight)
}

weighted_int_areas <- colSums(weighted_areas)

```

```{r dasymetric-time, echo=FALSE}

# The final weights used on each census block
stat_weights <- (weighted_int_areas/weighted_block_areas)/sum(weighted_int_areas/weighted_block_areas)

for(x in names(stat_weights)){
  stat_weights[x] <- polys[as.numeric(x)]$dog_prop * stat_weights[x]
}
print('Final answer!')
print(sum(stat_weights))


```
